\message{ !name(fdg_template.tex)}\documentclass[sigconf]{acmart}

\usepackage{booktabs} % For formal tables
\usepackage{epstopdf}

% Copyright
%\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}


\begin{document}

\message{ !name(fdg_template.tex) !offset(-3) }

\title{Attempting to Model a Commercial Choice-based Cinematic Interactive Narrative using Story Intention Graphs}
\author{Omitted for blind peer review}
\email{}
\affiliation{
  \institution{Omitted}
  \streetaddress{Omitted}
  \city{Omitted} 
  \state{} 
  \postcode{}
}

\begin{abstract}
This paper describes a planned study to use a model of the narrative
meaning to analyze a contemporary interactive digital narrative. The
primary research question is whether a model of the underlying story
can predict player emotional responses to key events within an
interactive narrative when provided with a player's earlier
choices. The game /The Wolf Among Us/ by Telltale Games was selected
due to both its critically acclaimed status and the fact that it has
both a strong, stable narrative and rich emotional content tied
directly to the story itself. The combination of sophisticated
narrative and high production values provides an opportunity to
understand how choices and branching stories operate within
contemporary digital narratives. The game uses choice menus with
natural language labels with simplistic accounting of inventory and
environments. The games uses bespoke hand authorship of complex
well-rounded characters and on dramatic voice and animated
performances, representing highly layered meanings and attracting
fervent fan communities.  We selected an existing computational model
of narrative for its ability to represent affective relationship
between story-values and the character goal networks that pursue them
(David Elson's Story Intention Graph and its text-focused annotation
tool, /Scheherazade/ \cite{Elson2012}) and plan to use annotated
adaptations of traversals to compare player experiences. This paper
describes the proposed efforts to pursue the aforementioned research
question: first, developing a repeatable methodology for annotating a
cinematic choice-based adventure using the selected narrative model
(SIG) and in particular associating events with story values and
characters. Second, recording a set of player's emotional experiences
while playing the game, and using the method developed in the first
effort to associate these responses with configurations of the story
content. Third, developing an algorithm whose input is the player's
traversal and whose output are potential places for emotional
response. Fourth, a second study validating the algorithm which is
conducted using episode 2 of /The Wolf Among Us/.
\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10010405.10010476.10011187.10011190</concept_id>
<concept_desc>Applied computing~Computer games</concept_desc>
<concept_significance>500</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Applied computing~Computer games}

% We no longer use \terms command
%\terms{Theory}

\keywords{computational narratology, story
  intention graph, interactive digital narratives}

\maketitle

\input{fdg_poster}

\bibliographystyle{ACM-Reference-Format}
\bibliography{../../../library} 
\end{document}

\message{ !name(fdg_template.tex) !offset(-99) }
